Comparative Analysis of Discount Rates Across Grocery Vendors

### 2024-11-21 paper/paper.qmd
Revise graph3

The bar plot(@fig-total) illustrates the average discount rate offered by each vendor. TandT and Walmart have the highest average discount rates, suggesting they tend to offer more substantial discounts than other vendors. On the other hand, NoFrills and Loblaws exhibit lower average discount rates, indicating a more conservative approach to discounting. The differences in average discount rates across vendors highlight varying discounting strategies that may be influenced by each vendor's pricing policies, target market, or competitive positioning.

@@

# Plot the data
ggplot(vendor_monthly_avg, aes(x = date, y = avg_discount_rate, color = vendor, group = vendor)) +
    geom_line() +
    labs(title = "Monthly Average Discount Rate by Vendor",
         x = "Date",
         y = "Average Discount Rate (%)") +
    theme_minimal() +
    scale_x_date(date_labels = "%Y-%m", date_breaks = "1 month") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
The line graph(@fig-monthly) tracks the monthly average discount rate for each vendor over the observed time period. We see considerable month-to-month variation in discount rates for certain vendors, particularly for SaveOnFoods, which shows a significant dip around July 2024, and Walmart, which reaches a peak around August 2024. The fluctuating patterns suggest that some vendors adjust their discount rates dynamically in response to factors such as seasonal demand, competitive pricing, or inventory levels. The chart provides a detailed view of how each vendor's discount strategy evolves over time, with some vendors maintaining more stable rates and others showing notable shifts across the months.
```{r}
#| label: fig-paradox
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false
#| fig-pos: "H"
ggplot(vendor_daily_avg, aes(x = as.Date(date), y = avg_discount_rate, color = vendor)) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(group = vendor), method = "lm", se = FALSE) +  # Within-vendor trend lines
  geom_smooth(data = vendor_daily_avg, aes(x = as.Date(date), y = avg_discount_rate), 
              method = "lm", color = "black", linetype = "dashed", se = FALSE) +  # Overall trend line
  labs(title = "Simpson's Paradox: Vendor Discount Trends",
       x = "Date",
       y = "Average Discount Rate",
       color = "Vendor") +
  theme_minimal()
```
The scatter plot(@fig-paradox) with trend lines illustrates the discount rates of various vendors over time, revealing an instance of Simpson's Paradox. While the overall trend may suggest a gradual increase or stable average discount rate, the individual vendor trends differ significantly. For example, some vendors exhibit a more consistent discount rate (such as Loblaws and Metro), while others, like TandT and Walmart, show more variability. This discrepancy between aggregate and individual trends highlights the importance of examining vendor-specific data to avoid drawing misleading conclusions from the aggregate trend alone.

>>>>

@fig-total illustrates the average discount rate offered by each vendor. TandT and Walmart have the highest average discount rates, suggesting they tend to offer more substantial discounts than other vendors. On the other hand, NoFrills and Loblaws exhibit lower average discount rates, indicating a more conservative approach to discounting. The differences in average discount rates across vendors highlight varying discounting strategies that may be influenced by each vendor's pricing policies, target market, or competitive positioning.

@@

# Plot the data
ggplot(vendor_monthly_avg, aes(x = date, y = avg_discount_rate, color = vendor, group = vendor)) +
    geom_line() +
    labs(title = "Monthly Average Discount Rate by Vendor",
         x = "Date",
         y = "Average Discount Rate (%)") +
    theme_minimal() +
    scale_x_date(date_labels = "%Y-%m", date_breaks = "1 month") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
@fig-monthly tracks the monthly average discount rate for each vendor over the observed time period. We see considerable month-to-month variation in discount rates for certain vendors, particularly for SaveOnFoods, which shows a significant dip around July 2024, and Walmart, which reaches a peak around August 2024. The fluctuating patterns suggest that some vendors adjust their discount rates dynamically in response to factors such as seasonal demand, competitive pricing, or inventory levels. The chart provides a detailed view of how each vendor's discount strategy evolves over time, with some vendors maintaining more stable rates and others showing notable shifts across the months.
```{r}
#| label: fig-paradox
#| fig-cap: "Trends in Vendor Discount Rates Over Time: Solid lines represent within-vendor trends, while the dashed black line indicates the overall trend across all vendors."
#| echo: false
#| warning: false
#| message: false
#| fig-pos: "H"
Sys.setlocale("LC_TIME", "C")
ggplot(vendor_daily_avg, aes(x = as.Date(date), y = avg_discount_rate, color = vendor)) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(group = vendor), method = "lm", se = FALSE) + # Within-vendor trend lines
  geom_smooth(data = vendor_daily_avg, aes(x = as.Date(date), y = avg_discount_rate),
              method = "lm", color = "black", linetype = "dashed", se = FALSE) + # Overall trend line
  scale_x_date(date_labels = "%b", # Format x-axis labels to show month names (e.g., Jan, Feb)
               date_breaks = "1 month") + # Add breaks at each month
  labs(title = "Simpson's Paradox: Vendor Discount Trends",
       x = "Month",
       y = "Average Discount Rate",
       color = "Vendor") +
  theme_minimal()
```
@fig-paradox with trend lines illustrates the discount rates of various vendors over time, revealing an instance of Simpson's Paradox. While the overall trend may suggest a gradual increase or stable average discount rate, the individual vendor trends differ significantly. For example, some vendors exhibit a more consistent discount rate (such as Loblaws and Metro), while others, like TandT and Walmart, show more variability. This discrepancy between aggregate and individual trends highlights the importance of examining vendor-specific data to avoid drawing misleading conclusions from the aggregate trend alone.
